---
title: 'Lab 3: Measuring Networks - Community Structure and Assortment'
author: "Tiernan Cahill"
date: "21/06/2021"
output:
  github_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Data and Required Packages
We will again be importing the graph objects to be used in this analysis from the serialized RDS format.

We will also need to load a few packages for this lab:

  - `dplyr` for wrangling data
  - `igraph` for modelling and visualizing our network
  - `igraphdata` for sample network data
  - `netseg` for alternative homophily and segregation measures
  
(*Note*: Community detection algorithms can be quite slow for larger networks, so for the purposes of quickly demonstrating multiple approaches, we will use a smaller network provided in the `igraphdata` package, which models friendships among faculty members at a university (Nepusz et al., 2008).)

```{r load, message = FALSE}
library(dplyr)
library(igraph)
library(netseg)

# Add faculty sample data to R environment (must have igraphdata installed)
data("UKfaculty", package = "igraphdata")

htag_graph <- readRDS("../data/htag_graph.RDS")
mention_graph <- readRDS("../data/mention_graph.RDS")
```

# Community Detection
There are a variety of different algorithms available as part of the `igraph` package for detecting the underlying community structure of networks. Most aim to maximise *modularity*, which is defined as the observed proportion of edges occurring within communities compared to what would be expected from a random graph. 

It should be noted that there is no one-size-fits-all approach to community detection. Not all algorithms will identify the optimal configuration, and modularity is known to be an imperfect measure of community structure (e.g., often biased against small communities in networks where large communities are present). All of these approaches also suffer from the limitation that the assigned community memberships are *exhaustive* and *exclusive* &mdash; every node belongs to a community, and to **only** one community.

```{r comm-detect, warning=FALSE}
# Newman and Girvan's (2004) modularity optimization algorithmm based on edge
# betweenness
cluster_edge_betweenness(UKfaculty, directed = T) -> comms.betweenness

# Pons and Latapy's (2005) modularity optimization algorithm based on random 
# walk distances
cluster_walktrap(UKfaculty) -> comms.randomwalk

# Clauset et al.'s (2004) hierarchical agglomerative method designed to improve
# performance on large, sparesly populated graphs (only works for undirected
# networks)
cluster_fast_greedy(as.undirected(UKfaculty)) -> comms.fastgreedy

# Let's try using the fast greedy method to detect communities in some 
# real-world social media networks
mention_comms <- mention_graph %>% 
  as.undirected() %>%
  cluster_fast_greedy()

htag_comms <- htag_graph %>%
  as.undirected() %>%
  cluster_fast_greedy()

# Blondel et al.'s (2008) multi-level modularity optimization algorithm (also 
# only works for undirected networks)
cluster_louvain(as.undirected(UKfaculty)) -> comms.louvain
```
All of the community detection algorithms included in `igraph` will output a `communities` object with a similar structure (some elements may depend on the particular algorithm used to generate the object). This object can be used and interpreted in a number of different ways, as shown below.
```{r comm-detect-output}
# Let's take a look at the output from one of the community detection algorithms
print(comms.louvain)

# How many communities are there?
length(comms.louvain)

# Compare this to the output from a different algorithm
length(comms.betweenness)

# How big is each community?
sizes(comms.louvain)

# Which community does each node belong to?
cbind(node = V(UKfaculty), 
      community = membership(comms.betweenness)) %>% 
  head(10)

# Try using the output of a different algorithm
cbind(node = V(UKfaculty), 
      community = membership(comms.randomwalk)) %>% 
  head(10)

# Which nodes are in each community?
communities(comms.randomwalk)

# What about a specific community
communities(comms.randomwalk)[[2]]
```
## Visualizing Community Detection
Once a community structure for a given network has been identified using a community detection algorithm, it can also be incorporated into visualizations of that network. For smaller networks, this can also be a useful way to compare the output of different algorithms to see which have face validity, or align with other attributes of the network.
```{r comm-detect-viz, fig.width = 10}
# Set an RNG seed so that we get consistent layouts
set.seed(9)
l = layout_with_fr(UKfaculty)

# Plot the whole network, colour-coded by the faculty members' departments
plot(UKfaculty,
     layout = l,
     edge.arrow.mode = 0,
     vertex.label = NA,
     vertex.size = 6,
     vertex.color = V(UKfaculty)$Group,
     main = "Faculty at a UK University Grouped by Department")

# Add the results of some community detection algorithms to the visualization
par(mfrow=c(1, 3))

plot(comms.betweenness, UKfaculty,
     layout = l,
     edge.arrow.mode = 0,
     vertex.label = NA,
     vertex.size = 6,
     main = "Edge Betweenness")

plot(comms.randomwalk, UKfaculty,
     layout = l,
     edge.arrow.mode = 0,
     vertex.label = NA,
     vertex.size = 6,
     main = "Random Walks")

plot(comms.louvain, UKfaculty,
     layout = l,
     edge.arrow.mode = 0,
     vertex.label = NA,
     vertex.size = 6,
     main = "Louvain Method")
```
We can also visualize the aggregation/division strategy used by the algorithm by plotting its output as either a *dendrogram* or a *hierarchical cluster*.
```{r comm-detect-viz2, fig.width = 10, fig.height = 12}
par(mfrow = c(2, 1))

# Plot as dendrograms
dendPlot(comms.fastgreedy, mode = "dendrogram", leaflab = "none", horiz = T,
       main = "Fast Greedy Method")

dendPlot(comms.randomwalk, mode = "dendrogram", leaflab = "none", horiz = T,
       main = "Random Walks")

# Plot as hierarchical clusters
dendPlot(comms.fastgreedy, mode = "hclust", labels = F,
       main = "Fast Greedy Method")

dendPlot(comms.randomwalk, mode = "hclust", labels = F,
       main = "Random Walks")
```

## Community Subgraphs
Once a community structure is established, this can be used to decompose a network into subgraphs. This can be particularly useful when attempting to break down a network with a *giant component* structure.

```{r comm-subgraphs}
# Extract Community #1 from the hashtag coincidence network
induced_subgraph(htag_graph, communities(htag_comms)[[12]]) -> htag_comm.12

# How big is this community?
vcount(htag_comm.12)

# Visualize the community
htag_comm.12 %>% 
  plot(vertex.shape = "none", 
       layout = layout_with_kk,
       vertex.label.color = "gray10",
       vertex.label.font = 2,
       edge.color = "gray85")
```

# Measures of Community
The modular structure of a network can be quantified and calculated in different ways, measuring the extent to which a network is clustered into discrete groups (i.e., *communities*), as well as the extent to which node-level attributes are related to the structure of these groups.

## Modularity
As mentioned above, *modularity* is a measure of the tendency of the nodes in a network to form discrete groups, defined as the proportion of edges occurring within a given community structure compared to a random graph:

$$Q = \frac{1}{2m} \left(A_{ij}-\frac{k_i k_j}{2m}\right)\delta\left(c_i,c_j\right)$$

Since modularity is calculated based on a pre-defined community structure, we can compare how this value varies depending on which community detection algorithm we use (and this can be a valuable metric when comparing the effectiveness of different algorithms). We can also define a community structure based on nominal node-level attributes, which can be especially useful when these attributes reflect real-world organizational units.

```{r calc-modularity}
# Calculate the modularity of a given community configuration
modularity(comms.betweenness)

# Now compare the modularity of different algorithmic approaches to 
# community configuration
cbind("modularity" = c(
        "edge betweenness" = modularity(comms.betweenness),
        "fast greedy method" = modularity(comms.fastgreedy),
        "louvain method" = modularity(comms.louvain),
        "random walks" = modularity(comms.randomwalk)))

# Modularity can also be calculated for arbitrary configurations (such as
# faculty department)
modularity(UKfaculty, membership = V(UKfaculty)$Group)
```

```{r calc-modularity-twitter}
# Let's try calculating modularity based on social media data!
cbind("modularity" = c(
  "user mentions" = modularity(mention_comms),
  "hashtag coincidence" = modularity(htag_comms)))
```

## Assortment
In contrast to modularity, *assortment* represents the tendency for nodes that have some defined attribute in common to be connected to each other. The assortment (or *assortativity*) coefficient is a modified form of the Pearson correlation intended to measure this tendency within a network:

$$r = \frac{\sum_{ij} \left(A_{ij}-k_ik_j/2m\right)x_ix_j}{\sum_{ij} \left(k_i\delta_{ij}-k_ik_j/2m\right)x_ix_j}$$
```{r calc-assortment}
# Calculate the assortment coefficient based on a nominal node-level attribute
assortativity_nominal(UKfaculty, V(UKfaculty)$Group, directed = T)

# Calculate the assortment coefficient based on degree centrality
assortativity_degree(UKfaculty, directed = T)
```

```{r calc-assortment-twitter}
# By degree
assortativity_degree(mention_graph, directed = T)

# The assortativity function doesn't like missing data, so we will have to replace
# this with a flag (noting that this *will* affect the results)
mention_graph.nomissing <- mention_graph
V(mention_graph.nomissing)$region[is.na(V(mention_graph.nomissing)$region)] <- 99L
V(mention_graph.nomissing)$followers[is.na(V(mention_graph.nomissing)$followers)] <- 0
V(mention_graph.nomissing)$verified[is.na(V(mention_graph.nomissing)$verified)] <- F

# By region
assortativity_nominal(mention_graph.nomissing, as.factor(V(mention_graph.nomissing)$region), directed = T)

# By followers
assortativity(mention_graph.nomissing, V(mention_graph.nomissing)$followers, directed = T)
```
### Alternative Homophily Measures
Assortment is one measure of *homophily*, or the general tendency of nodes in social networks to form connection with nodes with shared traits (i.e., "birds of a feather flock together"). While this is the primary measure of homophily included in the `igraph` package, additional measures of homophily (including some that provide group- or node-level calculations, such as Coleman's index) are available through the `netseg` package.

The most commonly-used of these is likely the *E/I Index*, which is defined as the ratio of between-group ($E$) to within-group ties($I$):

$$EI = \frac{E-I}{E+I}$$

Note that this means that a negative *E-I* index represents a tendency towards homophily (more within-group ties) and a postive index represents a tendency towards heterophily (more between-group ties).
```{r calc-homophily}
# The netseg package provides a number of additional measures of homophily and
# segregation
ei(UKfaculty, "Group")                      # Krackhard and Stern's E/I Index
freeman(as.undirected(UKfaculty), "Group")  # Freeman's segregation index
coleman(UKfaculty, "Group")                 # Coleman's homophily index
```

```{r calc-homophily-twitter}
# Let's try this on some real-world social media data!
ei(mention_graph, "verified")
freeman(as.undirected(mention_graph), "region")
coleman(mention_graph, "verified")

# Note that the functions from the netseg package can handle missing data, even
# for calculating the assortment coefficient
assort(mention_graph, "region")
```