---
title: 'Intro to Network Regression'
author: "Tiernan Cahill"
date: "23/06/2021"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Data and Required Packages
We will again be importing the graph objects to be used in this analysis from the serialized RDS format.

We will also need to load a few packages for this lab:

  - `readxl` for importing Excel data
  - `stringr` for parsing data in strings
  - `dplyr` for wrangling data
  -  `Matrix` for handling sparse matrices
  - `igraph` for modelling and visualizing our network
  - `statnet` for additional inferential analysis tools

```{r load, message = FALSE}
library(readxl)
library(stringr)
library(dplyr)
library(Matrix)
library(igraph)
library(statnet)
```

```{r data-import, message=FALSE}
# Import existing graph objects
mention_graph <- readRDS("../data/mention_graph.RDS")
htag_graph <- readRDS("../data/htag_graph.RDS")

# Import original dataset
lab <- read_excel("../data/lab.xlsx") %>% 
  rename_all(~str_replace_all(., "\\s+", "")) # Remove whitespace from variable names
```

```{r graph-gen}

# graph_gen(df, col) - Generate a sociomatrix based on a generic shared 
# attribute
graph_gen <- function (df, col) {
  
  # Create an affiliation matrix connecting users based on a shared attribute
  user_attr_matrix <- df %>%
    select(user = Author, attr = !! sym(col)) %>%
    table() %>%
    unclass() %>%
    Matrix()
  
  # Multiply the affiliation matrix by its transpose to create a sociomatrix
  shared_attr_matrix <- user_attr_matrix %*% t(user_attr_matrix)
  
  return(shared_attr_matrix)
}

# graph_gen_region(df) - Generate a sociomatrix based on shared region - alias
# for graph_gen()
graph_gen_region <- function (df) {
  return(graph_gen(df, col = "Region"))
}

# graph_gen_acctype(df) - Generate a sociomatrix based on shared account type -
# alias for graph_gen()
graph_gen_acctype <- function (df) { 
  return(graph_gen(df, col = "AccountType"))
}

# graph_gen_gender(df) - Generate a sociomatrix based on shared gender - alias
# for graph_gen()
graph_gen_gender <- function (df) {
  return(graph_gen(df, col = "Gender"))
}


# graph_gen_verified(df) - Generate a sociomatrix based on shared verified
# status - alias for graph_gen()
graph_gen_verified <- function (df) {
  return(graph_gen(df, col = "TwitterVerified"))
}
```

```{r graph-gen-multi}

# graph_gen_multi(df, col, sep) - Generate a sociomatrix based on a generic
# shared attribute that may contain multiple values
graph_gen_multi <- function (df, col, sep = ", ") {
  
  # Combine listed attributes from multiple tweets by the same user
  attr_list <- df %>%
    select(user = Author, attrs = !! sym(col)) %>%
    group_by(user) %>% mutate(attrs = paste(attrs, collapse = sep),
                              attrs = str_split(attrs, sep)) %>%
    rowwise() %>% mutate(attrs = list(unique(attrs)),
                         attrs = list(sort(attrs))) %>%
    distinct()
  
  # Convert the nested list to an affiliation matrix
  attr_matrix <- attr_list %>%
    unnest(attrs) %>%
    table() %>%
    unclass() %>%
    Matrix()
  
  # Multiply the affiliation matrix by its transpose to create a sociomatrix
  shared_attr_matrix <- attr_matrix %*% t(attr_matrix)
  return(shared_attr_matrix)
}
  
# graph_gen_htag(df) - Generate a sociomatrix based on shared hashtag usage -
# alias for graph_gen_multi()
graph_gen_htag <- function (df) {
  return(graph_gen_multi(df, col = "Hashtags"))
}

# Generate a sociomatrix based on shared interests - alias for graph_gen_multi()
graph_gen_interests <- function (df) {
  return(graph_gen_multi(df, col = "Interest"))
}
```