---
title: 'Intro to Network Regression'
author: "Tiernan Cahill"
date: "23/06/2021"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Data and Required Packages
We will again be importing the graph objects to be used in this analysis from the serialized RDS format.

We will also need to load a few packages for this lab:

  - `readxl` for importing Excel data
  - `dplyr` for wrangling data
  -  `Matrix` for handling sparse matrices
  - `igraph` for modelling and visualizing our network
  - `statnet` for additional inferential analysis tools

```{r load, message = FALSE}
library(readxl)
library(dplyr)
library(Matrix)
library(igraph)
library(statnet)
```

```{r data-import}
# Import existing graph objects
mention_graph <- readRDS("../data/mention_graph.RDS")
htag_graph <- readRDS("../data/htag_graph.RDS")

# Import original dataset
lab <- read_excel("../data/lab.xlsx") %>% 
  rename_all(~str_replace_all(., "\\s+", "")) # Remove whitespace from variable names
```

```{r graph-gen-htag}
# graph_gen_htag(df) - Generate a sociomatrix based on shared hashtag usage

graph_gen_htag <- function (df) {
  
  # Combine hashtag lists from multiple tweets by the same user
  htag_uses <- df %>%
    select(user = Author, htags = Hashtags) %>%
    group_by(user) %>% mutate(htags = paste(htags, collapse = ", "),
                              htags = str_split(htags, ", ")) %>%
    rowwise() %>% mutate(htags = list(unique(htags)),
                         htags = list(sort(htags))) %>%
    distinct()
  
  # Convert the nested list to a matrix
  htag_uses_matrix <- htag_uses %>%
    unnest(htags) %>%
    table() %>% unclass() %>% Matrix()
  
  # Multiply the affiliation matrix by its transpose to convert to a sociomatrix
  shared_htag_matrix <- htag_uses_matrix %*% t(htag_uses_matrix)
  
  return(shared_htag_matrix)
}
```

```{r graph-gen-region}
# graph_gen_region(df) - Generate a sociomatrix based on shared region
graph_gen_region <- function (df) {
  
  # Create an affiliation matrix connecting users to regions
  user_region_matrix <- lab %>%
    select(user = Author, region = Region) %>% 
    table() %>% 
    unclass() %>% 
    Matrix()
  
  shared_region_matrix <- user_region_matrix %*% t(user_region_matrix)
  
  return(shared_region_matrix)
}

```

```{r graph-gen}

# Generate a sociomatrix based on shared gender

# Generate a sociomatrix based on shared verified status

# Generate a sociomatrix based on shared interest

# Generate a sociomatrix based on shared account type


```