---
title: 'Intro to Network Regression'
author: "Tiernan Cahill"
date: "23/06/2021"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Data and Required Packages
We will again be importing the graph objects to be used in this analysis from the serialized RDS format.

We will also need to load a few packages for this lab:

  - `readxl` for importing Excel data
  - `stringr` for parsing data in strings
  - `dplyr` for wrangling data
  - `tidyr` for additional data wrangling
  -  `Matrix` for handling sparse matrices
  - `igraph` for modelling and visualizing our network
  - `statnet` for additional inferential analysis tools

```{r load, message = FALSE}
library(readxl)
# library(stringr)
# library(dplyr)
# library(tidyr)
# library(Matrix)
library(igraph)
library(statnet)

# Import helper functions
source("lab4_functions.R")
```

```{r data-import, message=FALSE}
# Import existing graph objects
mention_graph <- readRDS("../data/mention_graph.RDS")
htag_graph <- readRDS("../data/htag_graph.RDS")

# Import original dataset
lab <- read_excel("../data/lab.xlsx") %>% 
  rename_all(~str_replace_all(., "\\s+", "")) # Remove whitespace from variable names
```